Stateless Model Checking with Data-Race Preemption Points

### 2016-07-19 oopsla-16/evaluation.tex
buncha minor prose tweaks

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to \revisionx{SSS-MC-ICB (the fastest among the control experiments),} \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds.
\revisionx{Compared to SSS-MC-Shared-Mem (the best SSS-MC approach in the long term),} we ultimately concludes 10 CPU-hours with \revisionx{125\%} as many bugs in total \revisionx{when using Limited HB}.
\revision{Before the break-even point at 200 seconds,
\quicksand~lags behind \revisionx{SSS-MC-ICB} due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
Figure~\ref{fig:dowefindbugsfaster}(b) gives \quicksand~full credit for its inherent parallelism:
with a 10 core allocation, it outperforms SSS-MC
for any fixed budget of wall-clock time.}

@@

\revisionx{
To ensure that our corpus of P2 and Pintos bugs gives an unbiased comparison between \quicksand~and ICB,
we also counted the preemption bounds necessary for ICB to find each of its bugs.
Table~\ref{tab:icb-bounds} shows the distribution of these bounds,
which is consistent with the results of \cite{chess-icb}.
%justifies that our portrayal of ICB is fair.
%and shows no bias towards requiring high preemption bounds.
}

\begin{table}[h]
	\begin{center}
		\revisionx{
		\small
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf SSS-MC-ICB} & {\bf SSS-MC-Shared-Mem} \\
		\hline
		0	& 2	& 1	\\
		1	& 82	& 86	\\
		2	& 15	& 32	\\
		3	& 2	& 3	\\
		4+	& 0	& 0	\\
		\hline
		Total	& 101	& 122	\\
	\end{tabular}
		}
	\end{center}
	\label{tab:icb-bounds}
	\caption{\revisionx{Distribution of preemption bounds among bugs found by ICB \cite{chess-icb} control experiments. (Bound 0 means the bug was found by switching threads only on {\tt yield()}s.)}}
\end{table}

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{totalverifs-v2.pdf}
	\caption{Cumulative distribution of tests \quicksand~fully verified (\sect{\ref{sec:totalverif}}).
	%SSS-MC-Shared-Mem completed 1.37x as many tests, but 
	SSS-MC-Shared-Mem is unsound: 21 tests it completed hid bugs previously found in Figure~\ref{fig:dowefindbugsfaster}.
	%previously found (Figure~\ref{fig:dowefindbugsfaster}).
	%Some tests had no data-race candidates, and hence could also be verified by .}
	36 data-race-free tests were also soundly verified by QS-Sync-Only, SSS-MC-DPOR, and SSS-MC-ICB.}
	\label{fig:totalverif}
\end{figure}

>>>>

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to \revisionx{SSS-MC-ICB (the fastest among the control experiments),} \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds.
\revisionx{Compared to SSS-MC-Shared-Mem (the best SSS-MC approach in the long term), \quicksand's Limited HB version} ultimately concludes 10 CPU-hours with \revisionx{125\%} as many bugs in total.
\revision{Before the break-even point at 200 seconds,
\quicksand~lags behind \revisionx{SSS-MC-ICB} due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
Figure~\ref{fig:dowefindbugsfaster}(b) gives \quicksand~full credit for its inherent parallelism:
with a 10 core allocation, it outperforms SSS-MC
for any fixed budget of wall-clock time.}
\revisionx{After 1 hour of wall-clock time, tenfold \quicksand~performs 158\% as well as SSS-MC-ICB.}

@@

\revisionx{
To ensure that our corpus of P2 and Pintos bugs gives an unbiased comparison between \quicksand~and ICB,
we also counted the preemption bounds necessary for ICB to find each of its bugs.
Table~\ref{tab:icb-bounds} shows the distribution of these bounds,
which is consistent with the results of \cite{chess-icb}, showing no bias towards bugs that would be harder for ICB to find.
%justifies that our portrayal of ICB is fair.
%and shows no bias towards requiring high preemption bounds.
}

\begin{table}[h]
	\begin{center}
		\revisionx{
		\small
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf SSS-MC-ICB} & {\bf SSS-MC-Shared-Mem} \\
		\hline
		0	& 2	& 1	\\
		1	& 82	& 86	\\
		2	& 15	& 32	\\
		3	& 2	& 3	\\
		4+	& 0	& 0	\\
		\hline
		Total	& 101	& 122	\\
	\end{tabular}
		}
	\end{center}
	\caption{\revisionx{Distribution of preemption bounds among bugs found by ICB control experiments. (Bound 0 means the bug was found by switching threads only on {\tt yield()}s.)}}
	\label{tab:icb-bounds}
\end{table}

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{totalverifs-v2.pdf}
	\caption{Cumulative distribution of tests fully verified
	\revisionx{by QS-Limited-HB, QS-Pure-HB, and SSS-MC-Shared-Mem}
	(\sect{\ref{sec:totalverif}}).
	%SSS-MC-Shared-Mem completed 1.37x as many tests, but 
	%SSS-MC-Shared-Mem is unsound: 21 tests it completed hid bugs previously found in Figure~\ref{fig:dowefindbugsfaster}.
	%previously found (Figure~\ref{fig:dowefindbugsfaster}).
	%Some tests had no data-race candidates, and hence could also be verified by .}
	36 data-race-free tests were also soundly verified by QS-Sync-Only, SSS-MC-DPOR, and SSS-MC-ICB.}
	\label{fig:totalverif}
\end{figure}

### 2016-07-19 oopsla-16/related.tex
small parallel structure fix
buncha minor prose tweaks

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT \revisionx{targets} tests with impossibly large state spaces,
\revisionx{eschewing DPOR's depth-first search model to
%ensure it can
quickly sample a broad cross-section of large state spaces.
However, it proposes no alternate reduction algorithm to make up for its incompatibility with DPOR,
so is unsuitable for verification of medium-sized tests.}
%while for more approachable tests,
%its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
%
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB
%is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and can incorporate data-race PPs only when statically coded in advance}.
\revisionx{Our results in \sect{\ref{sec:eval-sssmc}} outperform both such configurations of ICB.}
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

\revisionx{{\bf Happens-Before.}}
%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive potential data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan}
%combine the lockset and happens-before analyses into a hybrid technique,
\revision{introduce Limited Happens-Before}, %(\sect{\ref{sec:overview-dr}})},
which
\revisionx{\quicksand~uses to achieve its best bug-finding result.
Other tools such as DJIT+ \cite{djit} and FastTrack \cite{fasttrack} opt for the precise Happens-Before relation first defined by Lamport \cite{lamport-clocks},
which produces no false positives but is prone to more false negatives as shown in \sect{\ref{sec:overview-dr}}.
FastTrack
%, the most recent development in this approach,
optimizes the representation of variable write vector clocks for performance,
which guarantees to detect at least the first race on each variable.
However, as we are interested in classifying data race candidates as {\em benign} or {\em buggy},
this optimization would be unsound for total verification if the first race on a variable were benign.
Hence, we implement precise HB with write vector clocks as defined by DJIT+.
Finally, \cite{predictive-dr} recently introduced the Causally-Precedes relation for a precise analysis which covers some common false negative cases
such as our example in Figure~\ref{fig:hb-example}(a).
}
%\revision{while \cite{predictive-dr} introduces the Causally-Precedes relation for a sound yet incomplete analysis (\sect{\ref{sec:overview-dr}})}.

>>>>

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT \revisionx{targets} tests with impossibly large state spaces,
\revisionx{eschewing DPOR's depth-first search model to
%ensure it can
instead sample broad cross-sections of large state spaces.
However, it proposes no alternate reduction algorithm to make up for its incompatibility with DPOR,
so is unsuitable for verification of medium-sized tests.}
%while for more approachable tests,
%its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
%
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB
%is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and can incorporate data-race PPs only when statically coded in advance}.
\revisionx{Our results in \sect{\ref{sec:eval-sssmc}} outperform both such configurations of ICB.}
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

\revisionx{{\bf Happens-Before.}}
%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive potential data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan}
%combine the lockset and happens-before analyses into a hybrid technique,
\revision{introduce Limited Happens-Before}, %(\sect{\ref{sec:overview-dr}})},
which
\revisionx{\quicksand~uses to achieve its best bug-finding result.
Other tools such as DJIT+ \cite{djit} and FastTrack \cite{fasttrack} opt for the precise Happens-Before relation first defined by Lamport \cite{lamport-clocks},
which produces no false positives but is prone to more false negatives as shown in \sect{\ref{sec:overview-dr}}.
FastTrack
%, the most recent development in this approach,
optimizes the representation of variable write clocks for performance,
which guarantees to detect at least the first race on each variable.
However, as we are interested in classifying data race candidates as {\em benign} or {\em buggy},
this optimization would be unsound for total verification if the first race on a variable were benign.
Hence, we implement precise HB with write vector clocks as defined by DJIT+.
Finally, \cite{predictive-dr} recently introduced the Causally-Precedes relation for a precise analysis which covers some common false negative cases
such as our example in Figure~\ref{fig:hb-example}(a).
}
%\revision{while \cite{predictive-dr} introduces the Causally-Precedes relation for a sound yet incomplete analysis (\sect{\ref{sec:overview-dr}})}.

### 2016-07-19 oopsla-16/soundness.tex
address the rest of reviewer D non-mandos
minor touch ups
buncha minor prose tweaks

We identify a particular class of false positive data-race candidates in which the associated memory was recycled by re-allocation between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behaviour.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 7 will be flagged as a \revision{potential}~data race.
We call this a {\em malloc-recycle data race \revision{candidiate}}.
To the human eye, this is obviously a false positive: reordering lines 4-7 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to an initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 6.

%However,
When limited to a single test execution, suppressing any data race \revision{candidate}~matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug},
in which the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Here, reordering lines 6-7 before line 3 will allow {\tt x} and {\tt x2} to race.

>>>>

We identify a particular class of false positive data-race candidates \revisionx{under Limited HB} in which the associated memory was recycled by re-allocation between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behaviour.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 7 will be flagged as a \revision{potential}~data race.
We call this a {\em malloc-recycle data race \revision{candidiate}}.
To the human eye, this is obviously a false positive: reordering lines 4-7 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to an initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 6.

%However,
When limited to a single test execution, suppressing any data race \revision{candidate}~matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug},
in which the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Here, reordering lines 6-7 before line 3 will allow {\tt x} and {\tt x2} to race.
\revisionx{Such collisions could be avoided with a hacked allocator which never recycles memory, but this could unacceptably impact performance in {\tt malloc}-heavy tests.}
